---
title: "Data Visualisation : Cameroon Alphabetisation/Work"
author: Janus Kuete
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r library}

library(readr)
library(tidyverse)
library(data.table)
library(ggrepel)
library(ggthemr)
library(maptools)
library(raster)
library(plyr)
library(ggplot2)
library(rgdal)
library(dplyr)
library(gganimate)
library(gifski)

```

## Statical anlysis of cameroun

http://www.dataforall.org/camsed/libraries/aspx/Home.aspx#


```{r Data Cleaning, echo = TRUE}
# Read the Datas from the csv file and fitting

datas <- read_csv("CamSED_2014_en.csv")

# remove the "#" and  remove Footnotes and groupby source to find the interesse and unusually , also bind Unit and indicator

textscript <- includeText("s.Rmd")
textserver <- includeText("StoryTelling/server.R")
textui <- includeText("StoryTelling/ui.R")



datas <- subset(as.tibble(lapply(datas, gsub, pattern = "#", replacement = "")), select = -Footnotes) %>%
  mutate(Values = as.numeric(`Data Value`)) %>%
  subset(select = -c(`Data Value`)) %>%
  transform(indicator__unit = paste(Indicator, Unit, sep = "__")) %>%
  subset(select = -c(Indicator, Unit)) %>%
  group_by(Source)

# seeing witch values are exploitable
tail(names(sort(table(datas$Time.Period))), 25) # 2014

tail(names(sort(table(datas$indicator__unit))), 25)

tail(names(sort(table(datas$Source))), 25)

# pivot Indicator

ins_datas <- datas %>%
  filter(grepl("CMR_INS_ECAM", Source)) %>%
  subset(select = -Source)
ins_datas <- ins_datas %>% pivot_wider(names_from = indicator__unit, values_from = Values)
```

## Employement Rate

```{r Employement Rate}

#Box config
box_config <- tibble::tribble(
  ~background, ~labelStatus,
  "red", "warning",
  "purple", "success",
  "green", "primary",
  "yellow", "danger",
  "fuchsia", "info"
)




# Employment Rate for Women and Men

ins_datas %>% 
  filter(Time.Period == "2014") %>%
  filter(Subgroup %in% c("Total", "Féminin", "Masculin")) %>%
  ggplot(aes(Area, `Taux d'emploi des personnes agées de 15-64 ans__Pour cent`, color = Subgroup, group = Subgroup ,label = `Taux d'emploi des personnes agées de 15-64 ans__Pour cent`))+
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  ylab("Employement Rate ") +
  theme(axis.text.x = element_text(
          colour = 'black', angle = 90,
          hjust = 0.5, vjust = 0.5),axis.title.x=element_blank()) 

# Employment Rate for Village and City

line <- function( liste)  {
  ggthemr_reset()
  ins_datas %>%
  filter(Time.Period == "2014") %>%
  filter(Subgroup %in% liste ) %>%
  ggplot(aes(Area, `Taux d'emploi des personnes agées de 15-64 ans__Pour cent`, color = Subgroup, group = Subgroup, label = `Taux d'emploi des personnes agées de 15-64 ans__Pour cent`)) +
  geom_line(size = 1) +
  geom_point(size = 1.5, colour = "black") +
  ggtitle(label = "Employement Rate  Rural or City ") +
  labs(fill = "percentage rate") +
  ylab("Employement Rate") +
  theme(axis.text.x = element_text(
    colour = "black", angle = 90,
    hjust = 0.5, vjust = 0.5
  ), axis.title.x = element_blank())
}
  
  
ggplotly(line(c("Féminin"))) 



```
- Die Männer haben ungefähr 10 % Mehr Chance ein Job zu haben 
- In Yaounde oder Douala ist es och größer 15 %
- In West ist es gleich
- Im Dorf gibt es mehr Chance zu arbeiten 
- Ist Landflucht ein Myth ??


## Primary School Rate

```{r Primary School}

# Primary School rate between women and men

bar <- function(variable , theme){  
  ggthemr(theme)
  ins_datas %>%
  filter(Time.Period == "2014") %>%
  filter(Subgroup %in% variable) %>%
  ggplot(aes(reorder(Area, `Taux net de scolarisation dans le primaire__Pour cent`), `Taux net de scolarisation dans le primaire__Pour cent`, fill = factor(Subgroup))) +
  ggtitle(label = "Primary School rate between women and men") +
  labs(fill = " Group") +
  ylab("Primary Scolarisation Rate ") +
  geom_col(position = "dodge", width = 0.5) +
  theme(axis.text.x = element_text(
    hjust = 0.5, vjust = 0.5
  ), axis.text.y = element_text(size = 7), axis.title.y = element_blank()) +
  coord_flip() }
  


```
-Mädschen und Junge haben die Gleiche Chance zu arbeien 
-Arme Regionen sind nicht favorisiert

## Region Alphabetisation


```{r Region Alphabetisation datacleaning}

# Alphabetisation in Region

Cameroon <- getData("GADM", country = "CMR", level = 1)

Cameroo <- spTransform(Cameroon, CRS("+init=EPSG:3119"))
Cameroon$NAME_1

Cameroon_df <- fortify(Cameroo)

# transform Ids to match with maps

Cameroon_df$id <- ifelse(Cameroon_df$id == 1, 28, Cameroon_df$id) # Adamawa
Cameroon_df$id <- ifelse(Cameroon_df$id == 2, 39, Cameroon_df$id) # Center
Cameroon_df$id <- ifelse(Cameroon_df$id == 3, 29, Cameroon_df$id) # East
Cameroon_df$id <- ifelse(Cameroon_df$id == 4, 30, Cameroon_df$id) # Far North
Cameroon_df$id <- ifelse(Cameroon_df$id == 5, 38, Cameroon_df$id) # Littoral
Cameroon_df$id <- ifelse(Cameroon_df$id == 6, 31, Cameroon_df$id) # North
Cameroon_df$id <- ifelse(Cameroon_df$id == 7, 32, Cameroon_df$id) # North-West
Cameroon_df$id <- ifelse(Cameroon_df$id == 8, 33, Cameroon_df$id) # West
Cameroon_df$id <- ifelse(Cameroon_df$id == 9, 34, Cameroon_df$id) # South
Cameroon_df$id <- ifelse(Cameroon_df$id == 10, 35, Cameroon_df$id) # South-west

# just select the needed ID for mapping

datastomap <- ins_datas %>%
  filter(Area.ID %in% c(28, 29, 30, 31, 32, 33, 34, 35, 38, 39)) %>%
  dplyr::rename("id" = "Area.ID")

mapdatas <- join(datastomap, Cameroon_df, by = "id")
```

```{r Region Alphabetisation Plot}

# removing unnecessary facets and plot(mapdatas)

g <- function(var , theme){  ggthemr(theme)
  mapdatas %>%
  filter(Subgroup == c(var)) %>%
  filter(Time.Period == "2014") %>%
  ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = `Taux d'alphabétisation des 15 à 49 ans__Pour cent`)) +
  ggtitle(label = "Alphabetisation in Regions") +
  labs(fill = "percentage rate", x = "Longitude", y = "Latitude") +
  theme(aspect.ratio = 1, axis.text.x = element_text(
    angle = 80,
    hjust = 0.5, vjust = 0.5
  )) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE))}
  



g('Féminin' , 'dust')

ggthemr_reset()
```
